An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-07-09 uai2017/paper.tex
Final version for UAI 2017 and arXiv

\subsection{BARKER (LOGISTIC) ACCEPTANCE FUNCTION}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\citep{Barker65}
function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
that it satisfies the condition in Lemma~\ref{lem:detailed_balance}.
It is slightly less efficient than the Metropolis test, since its
acceptance rate for vanishing likelihood difference is 0.5. However we
will see that its overall sample efficiency is much higher than the
earlier methods. See Appendix~\ref{app:why_barker} for details.

@@

Regarding our choice of the Logistic function, a test function $f(x)$ for
Metropolis-Hastings must satisfy Lemma~\ref{lem:detailed_balance}. In addition,
it must be monotone, bounded by $[0,1]$ and be such that $\lim_{x \to
-\infty}f(x) = 0$ and $\lim_{x \to \infty}f(x) = 1$. While many functions
satisfy this, including the classical test $f(x) = \min\{\exp(x), 1\}$, the
Logistic function is the \emph{unique} function in this class which is
anti-symmetric about 0.5, so it represents the (unique) CDF of a symmetric
random variable. Our method requires approximating this with the sum of a
Gaussian random variable (which is symmetric) and a correction. The Logistic CDF
$L$ and Gaussian CDF $\Phi$ are extremely close even without correction; more
precisely, the CDF error from the closest Gaussian CDF numerically satisfies
$\sup_x |L(x) - \Phi(x/1.7)| < 0.01$, i.e. the sup error between Logistic and
Gaussian CDFs is less than 1\%. With our correction we can make this error
orders of magnitude smaller.

>>>>

\subsection{BARKER (LOGISTIC) ACCEPTANCE FUNCTION}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\citep{Barker65}
function: $g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows
that it satisfies the condition in Lemma~\ref{lem:detailed_balance}.
It is slightly less efficient than the Metropolis test, since its
acceptance rate for vanishing likelihood difference is 0.5. However we
will see that its overall sample efficiency is much higher than the
earlier methods. See Appendix~\ref{app:why_barker} for additional discussion.

@@

Regarding our choice of the Logistic function, a test function $f(x)$ for
Metropolis-Hastings must satisfy Lemma~\ref{lem:detailed_balance}. In addition,
it must be monotone, bounded by $[0,1]$ and be such that $\lim_{x \to
-\infty}f(x) = 0$ and $\lim_{x \to \infty}f(x) = 1$. While many functions
satisfy this, including the classical test $f(x) = \min\{\exp(x), 1\}$, the
Logistic function is the \emph{unique} function in this class which is
anti-symmetric about 0.5, so it represents the (unique) CDF of a symmetric
random variable. Our method requires approximating this with the sum of a
Gaussian random variable (which is symmetric) and a correction. The Logistic CDF
$L$ and Gaussian CDF $\Phi$ are extremely close even without correction; more
precisely, the CDF error from the closest Gaussian CDF --- which we numerically
determined to have standard deviation approximately 1.7 --- satisfies $\sup_x
|L(x) - \Phi(x/1.7)| < 0.01$. Said another way, the error between the Logistic
and Gaussian CDFs is less than 1\%. With our correction we can make this error
orders of magnitude smaller.
